Predictive Quantization for MIMO-OFDM SVD Precoders using Reservoir Computing Framework

### 2019-04-13 paper/main.tex
Adding section 3 half

%Explain prediction and quantization steps here itself ! (succintly)
\section{Reservoir model for Predictive Quantization}
\label{section3}
\subsection{Vectorizing matrices in Stiefel Manifold}
%Focus on the fact that you don't actually pay a penalty
%The thing to quantize remains on the Stiefel Manifold
\subsection{Prediction and Quantization}
%Reservoir Pic
\section{Simulation Results}
\label{section4}
%Qtisn error with 1000 chan evols
%Qtisn error with 100 chan evols
%BER
%Achievable rate
\section{Conclusions and Future Work}
\label{section5}

>>>>

%Explain prediction and quantization steps here itself ! (succintly)
\section{Reservoir model for Predictive Quantization}
\label{section3}
\subsection{Vectorizing matrices in Stiefel Manifold}
The reservoir computing framework has inputs/outputs as vectors, and hence it is necessary to convert the precoder matrices to vectors. 
A precoding matrix $\mathcal{\bM} \in \text{St}(N_T,N_R)$ is a $N_T\times N_R$ matrix with the property $\mathcal{\bM}^H\mathcal{\bM}=\mathbb{I}_{N_R}$. 
For brevity sake, let us say that such a matrix $\mathcal{\bM}$ is a \'semi-unitary\' matrix.
A naive method of vectorizing a $N_T\times N_R$ semi-unitary matrix would be to take all it's $N_TN_R$ complex numbers individually and stack them onto a $2N_TN_R$ dimensional vector.
However, we can do this a little more efficiently, considering the $\mathcal{\bM}^H\mathcal{\bM}=\mathbb{I}_{N_R}$ property, which implies that each vector representing the columns of $\mathcal{\bM}$ is orthogonal to other columns.
For the first column, we take the $N_T$ complex numbers and stack them onto a $2N_T$ vector. 
For the second column, we take only the first $N_T-1$ complex numbers, since the last number can be determined by utilizing orthogonality with the first vector. 
For the third column, we take only the first $N_T-2$ complex numbers, and so on. 
Hence, we can vectorize the $N_T\times N_R$ $\mathcal{\bM}$ into a $2(N_T)+2(N_T-1)+\ldots+2(N_T-N_R+1)= 2N_TN_R-(N_R^2-N_R)$ dimensional vector. 
For any semi-unitary matrix $\mathcal{\bM}$, denote by $\mathcal{\bm}$ the vectorized representation.

Observe here that we do pay a dimensional penalty upon vectorizing the semi-unitary $\mathcal{\bM}$. 
$\mathcal{\bM} \in \text{St}(N_T,N_R)$, which is a $2N_TN_R-N_R^2$ dimensional manifold. 
We vectorize $\mathcal{\bM}$ to a $2N_TN_R-N_R^2+N_R$ dimensional vector, an overhead of $N_R$ dimensions, since we didn't exploit the fact that each column vector is a norm 1 vector as well.
It looks as if we could just take $N_T-1$ complex numbers when vectorizing the first column, instead of $N_T$, but this would lead to a non-unique representation, since there could then be infinite possibilities of the missing one complex number, such that the vector has a norm 1.
%Focus on the fact that you don't actually pay a penalty
%The thing to quantize remains on the Stiefel Manifold
\subsection{Prediction and Quantization}
The vector representation of semi-unitary matrices allow us to proceed with the discussion on reservoir computing framework for predictive quantization of SVD precoders. Consider the reservoir computing framework as in Fig. %Ref Fig. Reservoir Pic

The input to the reservoir $i$ corresponding to subcarrier $i$, is $\tilde{\bu}_{i,t-1}$, which is the vectorized representation of $\tilde{\bU}_{i,t-1}$. 
Let $D_{\text{in}}=2N_TN_R-N_R^2+N_R$ be the dimension of the input vector to the dimension.
Using a randomly initialized $D_{\text{in}}\times D_{\text{resv}}$ input coupler matrix $\bW^{\text{in}}_i$, the $D_{\text{in}}$ dimensional input is mapped to a $D_{\text{resv}}$ dimensional vector, where $D_{\text{resv}}$ is the dimension of the reservoir state vector $\br_{i,t}$. 
The reservoir state vector $\br_{i,t}$ is initialized with $D_{\text{resv}}$ zeros, and is updated via the following equation,
\begin{equation}
\br_{i,t}=\text{tanh}(\bA_i\br_{i,t-1}+\bW^{\text{in}}_i\tilde{\bu}_{i,t-1})
\label{res_upd}
\end{equation}
where $\bA_i$ is the adjacency matrix which captures the reservoir dynamics and correlates reservoir state across time, and tanh is applied element wise. 
Typical choices for $\bA_i$ have been erdos-renyi graphs with an upper bounded maximum eigen value \cite{mosleh2017brain,pathak2017using}.

The output of the reservoir is $\bp_{i,t}$, a vectorized representation of $\bP_{i,t}$. 
Hence, the output of the reservoir is also a $D_{\text{in}}$ dimensional vector ($D_{\text{out}}=D_{\text{in}}$). 
The output $\bp_{i,t}$ is obtained via matrix transformation of the reservoir state $\br_{i,t}$ via the $D_{\text{resv}}\times D_{\text{out}}$ output coupler matrix $\bW^{\text{out}}_i$, viz.
\begin{equation}
\bp_{i,t}=\bW^{\text{out}}_i\br_{i,t}
\label{pred_op}
\end{equation}

This completes the forward prediction algorithm to obtain $\bP_{i,t}$ using just one past observed $\bU_{i,t-1}$. That is,
\begin{equation}
\bP_{i,t}=f_P(\tilde{\bU}_{i,t-1})=\bW^{\text{out}}_i\text{tanh}(\bA_i\br_{i,t-1}+\bW^{\text{in}}_i\tilde{\bu}_{i,t-1})\footnote{For Brevity sake, $\bp_{i,t}$ and $\bP_{i,t}$ are treated to be one and the same}
\label{fwd_pred}
\end{equation}

However, a complete discussion of the prediction algorithm also involves a backward pass, involving the updation of output couplers $\bW^{\text{out}}_i$, which would follow next.

Before proceeding ahead, it is critical to detail out the quantization function $q_P(\bP_{i,t},\text{fb}_{i,t})$. 
The quantization method is similar to the one used in \cite{6891198,Gupt1905:Predictive,6545375}. 
Central to the quantization algorithm is the fact that tangent spaces local to a point in manifold, are actually linear vector spaces.
Given two points $\bX$ and $\bY$ in $\text{St}(N_T,NR)$, a lifting operation $\bT^Y_X=\texttt{lift}(\bX,\bY)$, $\texttt{lift}:\text{St}(N_T,NR)^2 \to \mathcal{T}_X \text{St}(N_T,NR)$ gives a tangent from X to Y, $\bT^Y_X \in \mathcal{T}_X \text{St}(N_T,NR)$, where $\mathcal{T}_X \text{St}(N_T,NR)$ is the local tangent space at $\bX$.
A corresponding retraction operation $Y=\texttt{retract}(X,\bT^Y_X)$gives back the manifold point obtained by traversing in the tangent direction.
In this work, the chosen lifting-retraction operations are the Cayley Exponentials, elaborated in \cite{DBLP:journals/corr/abs-1708-00045} and used in \cite{Gupt1905:Predictive}.
The Cayley exponential lifting operation maps two points in $\text{St}(N_T,N_R)$, to a $N_T\times N_T$ skew hermitian matrix representing the tangent from first point to the other.

The quantization algorithm exploits the linear vector space property of the tangent space and quantizes the local tangent space at the predicted precoder $\bP_{i,t}$.
The codebook for the local tangent space basically corresponds to a collection of codewords representing the different directions (tangents) in the local tangent space.
This codebook is also referred to as the base codebook for the quantization function.
The feedback from the receiver indicates the optimum direction in the base codebook, which the transmitter can choose to get closest to the actual value $\bU_{i,t}$.
With the optimum direction chosen, the next step is to determine how much to move in that particular direction, viz. the length of the chosen tangent direction.
For this, we adopt the strategy in \cite{6891198}, which
controls the magnitude of tangent steps, by having two codebooks
$T^{\mathcal{C}}_p,T^{\mathcal{C}}_m$ of different spreads, but same
$2^{\mathcal{B}-1}$ base vectors from the base codebook. 
The base vectors correspond to $2^{\mathcal{B}-1}$ quantized directions, and by spread of codewords, we mean the length of base vectors considered, with $T^{\mathcal{C}}_p$ having tangents of higher length than $T^{\mathcal{C}}_m$. 
The two codebooks are concatenated to form a $2^\mathcal{B}$ length codebook $T^\mathcal{C}$. 
The receiver finds the optimal index $c_n$ in $T^\mathcal{C}$ by comparing the chordal distance metric ($d_s$) of Stiefel Manifold \cite{Gupt1905:Predictive,6891198}, of each codeword to the actual precoder $\bU_{i,t}$ obtained from the SVD of the channel matrix, using \eqref{eqn:cn1}. The receiver then feeds back $c_n$ to the transmitter using $\mathcal{B}$ bits. The transmitter uses the fed back $c_n$ and \eqref{eqn:cn2} to calculate $\tilde{\bU}_{i,t}$,
\begin{equation}
\label{eqn:cn1}
c_n \gets \text{argmin}_{i \in 2^\mathcal{B}}\left (d_s\left (\bU_{k,l},\text{retract}(\bP_{k,l},T^\mathcal{C}[i])\right ) \right)
\end{equation}
\begin{equation}
\tilde{\bU}_{k,l}=\text{retract}(\bP_{k,l},T^\mathcal{C}[c_n])
\label{eqn:cn2}
\end{equation}

% \textcolor{blue}
{We now describe the algorithm used to control the spread of the codebooks, as discussed in \cite{6891198}. The base codebook ($\mathcal{T}^B$) consists of
  $2^{\mathcal{B}-1}$ matrices belonging to the local tangent space at
  $\bP_{k,l}$, which represent the $2^{\mathcal{B}-1}$ quantized directions. The two codebooks,
  $T^{\mathcal{C}}_p,T^{\mathcal{C}}_m$, have the same set of base
  vectors, but different spreads $g^{s_p}$ and $g^{s_m}$, where $g$ is
  the growth factor and $s_p,s_m$ control the spread of the two
  codebooks. Depending on whether $c_n \geq 2^{\mathcal{B}-1}$,
  i.e. whether the optimum codeword is in $T^{\mathcal{C}}_p$ or
  $T^{\mathcal{C}}_m$, the scale parameter $s[k]$, which in turn
  controls values of $s_p, s_m$ is updated in the following manner,
$$s_p=g^{\text{min}(s[k-1]+1,0)}, s_m = g^{s[k-1]-1}$$
$$
s[k]=
\begin{cases}
\text{min}(s[k-1]+1,0), \text{ for }c_n \in T^{\mathcal{C}}_p\\
s[k-1]-1, \text{otherwise}\\
\end{cases}
$$
with $s[0]=0$. Intuitively, the algorithm reduces/increases the spread
of the codebook till the operation of reduction/increase is no longer
beneficial, i.e. the optimum codeword lies in the higher/lower spread
codebook instead. The scheme we use for obtaining the quantized
directions, however, differs slightly from the one presented in
\cite{6891198}. As discussed in Section \ref{section3}, the Cayley
type lifting map takes inputs from the Stiefel manifold and maps them
to $N_T \times N_T$ skew Hermitian matrices. This allows for vector
quantization of the tangent space, since a skew Hermitian matrix can
be easily converted to a vector, by taking the upper triangular
entries and stacking them in a vector. Vector quantization for the
base codebook, $\mathcal{T}^B$ allows for isotropic $2^{\mathcal{B}-1}$ quantized
directions for the tangent space local to $\bP_{k,l}$. \cite{6891198}
proposed a randomly initialized $\mathcal{T}^B$ which does not guarantee
isotropic $2^{\mathcal{B}-1}$ quantized directions, which can
potentially hurt the performance of the system in the cases when
random initialization gives directions which are close to each
other. More details on codebook construction are provided in Section
\ref{section5}. This concludes the proposed predictive quantization
algorithm.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% More information needed about the adjacency
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Simulation Results}
\label{section4}
%Qtisn error with 1000 chan evols
%Qtisn error with 100 chan evols
%BER
%Achievable rat
\section{Conclusions and Future Work}
\label{section5}
